---
title: "AIDS Drug"
author: "Crystal Escobar"
date: "May 31, 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(dplyr)
avast <-read.csv(file="avastivironal.csv",header=TRUE)
head(avast)
females<-filter(avast,sex=="F")
males<-filter(avast, sex=="M")
avastIV = filter(avast, Ivdrug=="user")
avastLBGT= filter(avast, LBGT=="yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides (abacavirinol) is under test in clinical trials. Analyze the effectiveness of this drug across the board. Consider population subgroups by gender, drug use, and sexual orientation.

The first question is this: Does the AIDS drug, abacavirinol have any effect on whether one contracts AIDS or not?  

$H_0:$ The AIDS drug, abacavirinol has no effect on whether one contracts AIDS or not.  
$H_1:$ The AIDS drug, abacavirinol has an effect on whether one contracts AIDS or not.  

The remaining questions are: Does the AIDS drug, abacavirinol have a different effect based on gender, drug use, or sexual orientation?    

$H_0$ Gender has no effect on whether abacavirinol offers any protection against contracting AIDS.    
$H_1$ Gender has an effect on whether abacavirinol offers any protection against contracting AIDS.  

$H_0$ Drug use has no effect on whether abacavirinol offers any protection against contracting AIDS.   
$H_1$ Drug use has an effect on whether abacavirinol offers any protection against contracting AIDS. 

$H_0$ Sexual orientation has no effect on whether abacavirinol offers any protection against contracting AIDS.   
$H_1$ Sexual orientation has an effect on whether abacavirinol offers any protection against contracting AIDS.  


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

For each hypothesis test, the graphical summary will be a bar chart, the numerical summary will be a cross tabulation, and a chi-squared test will be used for the inferential results.  

#Results- General Effectiveness
Is the vaccine effective in the general population?

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


## Descriptive Results

We use a barchart, cross tabulation, and chi-squared test to show the effectiveness of the AIDS drug, abacavirinol on the general population.

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

We use a barchart to show the effectiveness of the AIDS drug, Abacavirinol. The chart shows that when on treatment the general population is less likely to contract the AIDS virus. 

```{r echo=FALSE}
barchartGC(~infected+treatment, data=avast, type= 'percent', xlab= "General Population")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical Summary is a cross tabulation.  

```{r echo=FALSE}
gentable = xtabs(~infected+treatment, data=avast)
print(gentable)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

A chi-squared test is used to determine the inferential results.  

```{r echo=FALSE}
chisq.test(gentable)
```

The hypothesis was that the effectiveness of the vaccine on the general population is related to whether one contracted the AIDS virus. We used a chi squared of independence test to test the hypothesis. The P-value is less than 0.05, which means our conclusion is that we reject the null hypothesis, and accept the alternative hypothesis. We have evidence that the vaccine is effective and is related to whether one contracts the AIDS virus.

#Results- Gender Effects
Does Gender make a difference in vaccine effectiveness?

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

We use a barchart to show the effectiveness of the AIDS drug, Abacavirinol. The charts show that the vaccine increases the chances of females contracting the AIDS virus and decreases the chances for males.

```{r echo=FALSE}
barchartGC(~infected+treatment, data=females, type = "percent", xlab= "Females")
```

```{r echo=FALSE}
barchartGC(~infected+treatment, data=males, type = "percent", xlab= "Males")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical Summary is a cross tabulation.  

```{r echo=FALSE}
gentableF = xtabs(~infected+treatment, data=females)
print(gentableF)
```

```{r echo=FALSE}
gentableM = xtabs(~infected+treatment, data=males)
print(gentableM)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

A chi-squared test is used to determine the inferential results.    

```{r echo=FALSE}
chisq.test(gentableF)
fisher.test(gentableF)
```

```{r echo=FALSE}
chisq.test(gentableM)
fisher.test(gentableM)
```

The hypothesis was that the effectiveness of the vaccine based on gender is related to whether one contracted the AIDS virus. We used a chi squared of independence test to test the hypothesis. We got a P-value less than .05 for females, which means our conclusion is that we reject the null hypothesis and accept the alternative hypothesis. We have evidence that the vaccine increases susceptibility to the AIDS virus in females. We got a P-value less than .05 for males, which means our conclusion is that we reject the null hypothesis and accept the alternative hypothesis. We have evidence that the vaccine decreases susceptbility in males and if on treatment the chances of contracting the virus are less likely.  

#Results- IV Drug Use Effects
Does IV drug use make a difference in vaccine effectiveness?  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

We use a barchart to show the effectiveness of the AIDS drug, Abacavirinol. The chart shows that the vaccine increases the chances of IV users contracting the AIDS virus.

```{r echo=FALSE}
barchartGC(~infected+treatment, data=avastIV, type = "percent", xlab= "IV Users")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical Summary is a cross tabulation.  

```{r echo=FALSE}
gentableIV = xtabs(~infected+treatment, data=avastIV)
print(gentableIV)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

A chi-squared test is used to determine the inferential results.     

```{r echo=FALSE}
chisq.test(gentableIV)
fisher.test(gentableIV)
```

The hypothesis was that the effectiveness of the vaccine based on IV users is related to whether one contracted the AIDS virus. We used a chi squared of independence test to test the hypothesis. We got a P-value less than .05 which means our conclusion is that we reject the null hypothesis and accept the alternative hypothesis. We have evidence that the vaccine increases susceptibilty in IV users, and if on treatment you are more likely to contract the virus. 


#Results- Sexual Orientation Effects
Does sexual orientation make a difference in vaccine effectiveness?  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

We use a barchart to show the effectiveness of the AIDS drug, Abacavirinol. The chart shows that the vaccine decreases the chances of the LBGT contracting the AIDS virus.

```{r echo=FALSE}
barchartGC(~infected+treatment, data=avastLBGT, type = "percent", xlab= "LBGT")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical Summary is a cross tabulation.    

```{r echo=FALSE}
gentableLBGT = xtabs(~infected+treatment, data=avastLBGT)
print(gentableLBGT)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

A chi-squared test is used to determine the inferential results.   

```{r echo=FALSE}
chisq.test(gentableLBGT)
fisher.test(gentableLBGT)
```

The hypothesis was that the effectiveness of the vaccine based on LBGT is related to whether one contracted the AIDS virus. We used a chi squared of independence test to test the hypothesis. We got a P-value less than .05 which means our conclusion is that we reject the null hypothesis and accept the alternative hypothesis. We have evidence that the vaccine decreases susceptability and if on treatment the chances of contracting the virus is less likely.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The overall hypothesis was that the effectiveness of the vaccine on the general population, gender, IV drug users, and sexual orientation is related to whether one contracted the AIDS virus. We used a chi squared of independence test to test the hypothesis. The P-value for each category is less than .05, which means our conclusion is that we reject the null hypothesis, and accept the alternative hypothesis. We have evidence that the vaccine is effective and is related to whether one contracts the AIDS virus. Our results show that the vaccine increases susceptibility in females and IV drug users. It also shows that the vaccine decreases susceptibility in males, LBGT, and the general population. 
